---
title: "Martino"
author: "Martino Armanini"
date: "26/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# load libraries
library(tidyverse)
library(vroom)
library(janitor)
library(sf)
library(here)
library(extrafont)
library(ggtext)
library(lubridate)
library(urbnmapr)
library(readxl)
library(stringr)
library(ggnewscale)
library(gganimate)
library(animation)
library(transformr)
library(skimr)

loadfonts(device="win")
```

```{r}
# load "firsts" data and clean column names
firsts <- read_csv(here("../data/firsts_augmented.csv")) %>% 
  clean_names()
```
# Glimpse and Skim the data
```{r}
glimpse(firsts)

skim(firsts)

```

From this first analysis we see we're having quite a significant amount of missing values.

Let's load the other data set before modifying anything from here

#Loading population

```{r}
# load population data from 1790-2010 and clean names
# source: https://conservancy.umn.edu/handle/11299/181605
pop <- read_xlsx(here("../data/county2010_hist_pops.xlsx"), sheet = "c2010_hist_pops") %>% 
  clean_names()

```

Data Discovery

# Time series
```{r}

firsts <- firsts %>% 
  mutate(year_2 = cut(year, 
                      breaks = c(min(year)-1, c(seq(1790, 2020, 10))), 
                      labels = c(seq(1790, 2020, 10)))) %>% 
  mutate(year_2 = as.integer(levels(year_2))[year_2]) %>% 
  mutate(year_full = year) %>%
  mutate(year = year_2) %>% 
  select(-year_2)


#Changing the format of year column
firsts$year  <- as.Date(as.character(firsts$year), "%Y")
firsts$year <- year(firsts$year)



str(firsts)
#date_test <- c_d_clean %>% 
#  mutate(day1 = day(date_rep))




time <- firsts %>% 
  group_by(year) %>% 
  count(year) %>% 
  rename("n_first" = "n") %>% 
  mutate(growth = ((42-9)/9)*100)

time$growth <- paste(round(time$growth), "%", sep="")

ggplot(time, aes(x = year, y = n_first)) +
  geom_line() +
  geom_point() +
  geom_smooth(method= lm, se =F )
#ADD Percentages

label1 <- "Over the years, \n the first times by decade \n have increased by 367%"

time_firsts <- time %>%
 ggplot(aes(x = year, y = n_first, group = 1)) +
 geom_line(colour = "aquamarine4", size = 1.2, lty = 2)+
 geom_point(pch=16, size = 3, colour = "salmon")+
 geom_smooth(method = lm, se = FALSE, colour = "Royalblue4") +
 theme_bw()+
 labs(colour = "stoppage",
 title = "A positive growth of first times for African Americans",
 subtitle = "The number of first times for the Black community in the US",
 caption = "Source: XXXXX",
 x = "",
 y = "Number of firts") +
  geom_text(
    data = data.frame(x = 1875, y = 55, label = label1),
    aes(x = x, y = y, label = label1),
    colour="Royalblue4",
    family="Times New Roman",
    size = 5,
    hjust = 0.5,
    lineheight = .8,
    inherit.aes = FALSE)+
#  geom_text(
#  data = data.frame(x = 26.6, y = 880, label = label),
#  aes(x = x, y = y, label = label4),
#  colour="salmon",
#  family="Lato",
#  size = 5,
#  hjust = 0.5,
#  lineheight = .8,
#  inherit.aes = FALSE)+
#  geom_text(
#  data = data.frame(x = 5.5, y = 410, label = label),
#  aes(x = x, y = y, label = label5),
#  colour="darkgreen",
#  family="Lato",
#  size = 5,
#  hjust = 0.5,
#  lineheight = .8,
#  inherit.aes = FALSE)+
#  geom_text(
#  data = data.frame(x = 23.4, y = 880, label = label),
#  aes(x = x, y = y, label = label6),
#  colour="darkgreen",
#  family="Lato",
#  size = 5,
#  hjust = 0.5,
#  lineheight = .8,
#  inherit.aes = FALSE)+ # geom_curve(
# # data = data.frame(x = 6000, y = 5.3, xend = 6500, yend = 7.5),
# # mapping = aes(x = x, y = y, xend = xend, yend = yend),
# # colour = "grey15",
# # size = 0.5,
# # curvature = 0.25,
# # arrow = arrow(length = unit(2, "mm"), type = "closed"),
# # inherit.aes = FALSE)+
 theme_classic()+
 theme(legend.position = "none",
 plot.title = element_text(size = 20, face = "bold", colour = "aquamarine4"),
 plot.subtitle = element_text(size = 15),
 axis.title = element_text(size = 14),
 axis.text.x = element_text(angle = 0,  hjust = 0.5, size = 14),
 axis.text.y = element_text(angle = 0, hjust = 0.5, size = 14))


time_firsts


```


## Most firsts per state and distribution of firts
```{r}

label2 <- "New York appears to be \n the most forward thinking state \n in the US with 8% \n of the total firsts"
my_colours <- c("grey80", "dodgerblue1")

states <- firsts %>% 
  filter(!is.na(state)) %>% 
  count(state) %>% 
  mutate(tot  = sum(n), pct = n/sum(n)*100,
         NY = ifelse(`state` == "New York", TRUE, FALSE)) %>%
  filter(pct > 2) %>% 
  arrange(desc(n))

states$pct <- paste(round(states$pct), "%", sep="")
  

states <- states[-c(28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43),]

ggplot(states, aes(n)) + 
  geom_histogram()

ggplot(states, aes(x = n, y = reorder(state, n))) +
  geom_col() +
  theme_classic()

ggplot(states, aes(x =  n, 
                     y = reorder(state, n), fill = NY)) + # reorder bars by beer servings
  geom_col() + 
  scale_fill_manual(values = my_colours) + 
  # geom_text(
  #   aes(label = n, x = n - .25),
  #   colour = "red4",
  #   size = 5,
  #   nudge_x = 400,
  #   nudge_y = 0,
  #   hjust = 0.45,
  #   family="Lato") +
  geom_text(
    aes(label = pct),
    colour = "green4",
    size = 6,
    nudge_x = 1,
    nudge_y = 0,
    hjust = 0.5,
    family="Lato")+
  geom_text(
    data = data.frame(x = 25, y = 9.5, label = label2),
    aes(x = x, y = y, label = label2),
    colour="dodgerblue1",
    family="Lato",
    size = 6,
    hjust = 0.5,
    lineheight = .8,
    inherit.aes = FALSE)+
  # geom_text(
  #   data = data.frame(x = 2500, y = 0.75, label = label2),
  #   aes(x = x, y = y, label = label2),
  #   colour="green4",
  #   family="Lato",
  #   size = 4.5,
  #   hjust = 0.5,
  #   lineheight = .8,
  #   inherit.aes = FALSE)+
  geom_curve(
    data = data.frame(x = 25, y = 10, xend = 30, yend = 19.2),
    mapping = aes(x = x, y = y, xend = xend, yend = yend),
    colour = "grey15",
    size = 0.5,
    curvature = 0.25,
    arrow = arrow(length = unit(2, "mm"), type = "closed"),
    inherit.aes = FALSE) +
  labs(colour = "stoppage",
       title = "The most liberal state takes the highest number of firsts",
       subtitle = "Over time, New York State has outnumbered the other states",
       caption = "Source: Data.Police.UK",
       x = "Number of Stops and Searches in September 2020",
       y = "") +
  theme_classic()+
  theme(legend.position = "none",
        plot.title = element_text(size = 23, face = "bold", colour = "dodgerblue1"),
        plot.subtitle = element_text(size = 15),
        axis.title = element_text(size = 14),
        axis.text.x = element_text(angle = 0, hjust = 1, size = 14),
        axis.text.y = element_text(angle = 0, hjust = 1, size = 14))




```

Interesting stuff, we can see that there is a stark difference between state 36 (New York State) and state 27(Minnesota) we'll see later how these developed over time! (Why is Mexico in there?)

## Gender and Category analysis

```{r}

my_colours2 <- c("skyblue1", "palevioletred1")

## Gender count
gender <- firsts %>% 
  filter(!is.na(gender)) %>% 
  count(gender) %>% 
  mutate(tot = sum(n), pct = n/tot*100, 
         female = ifelse(`gender` == "female", TRUE, FALSE)) %>% 
  arrange(desc(n))

gender$pct <- paste(round(gender$pct), "%", sep="")

ggplot(gender, aes(x = gender, y = n, fill = female)) + 
  geom_col() +
  scale_fill_manual(values = my_colours2) +
   geom_text(
    aes(label = pct),
    colour = "Navyblue",
    size = 12,
    nudge_x = 0,
    nudge_y = 12,
    hjust = 0.5,
    family = "Lato") +
  theme_classic()+
    labs(colour = "gender",
       title = "Sexism is part of African American History",
       subtitle = "Males have been first more than 60% of the time",
       caption = "Source: XXXX",
       x = "Gender",
       y = "Number of firsts by gender") +
  theme(legend.position = "none",
        plot.title = element_text(size = 30, face = "bold", colour = "Navyblue"),
        plot.subtitle = element_text(size = 15),
        axis.title = element_text(size = 14),
        axis.text.x = element_text(angle = 0, hjust = 0.5, size = 14),
        axis.text.y = element_text(angle = 0, hjust = 1, size = 14))
  


## Category count
category <- firsts %>% 
  filter(!is.na(category)) %>% 
  count(category) %>% 
  mutate(tot = sum(n), pct = n/tot*100) %>% 
  arrange(desc(n))
  

ggplot(category, aes(x = n, y = reorder(category, -n), fill = n)) +
  geom_col() +
  scale_fill_gradient(low = "red3", high = "navy") +
   theme_classic()+
    labs(colour = "gender",
       title = "Law and Religion are the most conservative areas",
       subtitle = "First times for African American by category",
       caption = "Source: XXXX",
       x = "",
       y = "") +
  theme(legend.position = "none",
        plot.title = element_text(size = 30, face = "bold", colour = "Navyblue"),
        plot.subtitle = element_text(size = 15),
        axis.title = element_text(size = 14),
        axis.text.x = element_text(angle = 0, hjust = 0.5, size = 14),
        axis.text.y = element_text(angle = 0, hjust = 1, size = 14))
  
# ADD PERCENTAGES


```


# US Map
```{r}

# load states shapefile
states_sf <- get_urbn_map("states", sf = TRUE)
states_sf$geometry

# load counties shapefile
counties_sf <- get_urbn_map("counties", sf = TRUE)
counties_sf$geometry


# transfrom geometry to 4326, or pairs of latitude/longitude numbers
states_sf <-  states_sf %>% 
  st_transform(4326) # transfrom to WGS84, latitude/longitude
states_sf$geometry

# transfrom geometry to 4326, or pairs of latitude/longitude numbers
counties_sf <-  counties_sf %>% 
  st_transform(4326) # transfrom to WGS84, latitude/longitude
counties_sf$geometry


# join counties_sf and population data
counties_pop_sf <- counties_sf %>% 
  left_join(pop, by = c("county_fips" = "geoid10"))


# convert firsts to a sf object
firsts_sf <- firsts %>% 
  drop_na(lng, lat, gender) %>% 
  filter(country == "United States of America") %>% 
  # filter irrelevant lng/lat
  filter(lng > -140) %>% 
  st_as_sf(coords = c('lng', 'lat'), 
           crs = st_crs(states_sf))
firsts_sf$geometry

# convert firsts to a sf object with jitter on coordinates
# jitter points such that they are better visible
set.seed(100)
firsts_sf_jitter <- firsts %>% 
  drop_na(lng, lat, gender) %>% 
  mutate(lng = jitter(lng, amount = 1),
         lat = jitter(lat, amount = 1)) %>% 
  filter(country == "United States of America") %>% 
  # filter irrelevant lng/lat
  filter(lng > -140) %>% 
  st_as_sf(coords = c('lng', 'lat'), 
           crs = st_crs(states_sf))
firsts_sf_jitter$geometry



ggplot() +
  # draw polygons from states shapefile
  geom_sf(data = states_sf, fill = "grey99", color = "black")+
  # add points from firsts shapefile
  geom_sf(
    data = firsts_sf, aes(fill = gender), size = 2, shape = 21,
    show.legend = TRUE
  ) + 
  scale_fill_manual(values = c("palevioletred1", "skyblue1")) +
  theme_minimal()+
  coord_sf(datum = NA) + #remove coordinates
  labs(title = "Most achievers were born in the East!") +
  # hrbrthemes::theme_ft_rc(grid="", strip_text_face = "bold") +
  theme(axis.text = element_blank()) +
  # theme(strip.text = element_text(color = "white"))+
  NULL
```

```{r}

```




