---
title: "Noor"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(cluster)
library(Hmisc)
library(factoextra)
library(purrr)
library(gridExtra)
library('ggplot2')
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(dplyr)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(kableExtra)
library(vroom)
library(leaflet)
library(moderndive)
library(car)
library(huxtable)
library(caret)
library(sf) # for geospatial visualisation
library(rsample)
library(extrafont)
library(vroom)
library(ggtext)
library(gapminder)
library(ggrepel)
library(patchwork)
library(gghighlight)
library(skimr)
library(ggplot2)
library(lubridate)
library(plotly)
```


load the data
```{r}
library(readr)
firsts<- read_csv("~/GitHub/am10.sg10/data/firsts_augmented.csv")
#glimpse(firsts)

firsts <- firsts %>% filter( country == "United States of America")
describe(firsts)
```
Load map data and join to firsts data
```{r}
library(urbnmapr)

counties_sf <- get_urbn_map("counties", sf = TRUE)

class(counties_sf)
counties_sf$geometry

#counties_sf %>% 
#  ggplot(aes()) +
#  geom_sf(fill = "grey70", color = "#ffffff")
  
firsts_joined<- left_join(counties_sf, firsts, by = c("county_fips" =  "FIPS_state"))
#firsts_sf = subset(firsts_sf, select = -c(state_name , fips_class,county_name,state_fips ,state_abbv))

firsts_joined_sf <-  st_as_sf(firsts_joined, 
                              coords=c('lng', 'lat'), 
                              crs = 4326)  



firsts_joined_sf_statecount <- firsts_joined_sf%>% group_by(state_name)%>% mutate(countrow = n())

#skimr::skim(firsts_joined_sf)


```



# plot for challenge 1 

```{r}
plot1 <-   
    ggplot(data=firsts_joined_sf_statecount )+
    geom_sf( fill = "grey70", colour = "white")+
    geom_sf(
      aes(fill = countrow)
    )+
    geom_sf_text(data = get_urbn_labels(map = "states", sf = TRUE), 
                 aes(label = state_abbv), 
                 size = 5,
                 family="Lato") +
    theme_void()+
  #facet_wrap(~gender)+
    labs(
      title = "Number of firsts per state",
      #subtitle = "US Presidential Election 2020",
      #caption = "Data Source: New York Times",
      x=NULL, y=NULL
    )+
    theme(text=element_text(size=12, family="Lato"))+
    theme(plot.title.position = "plot")+
    NULL
plot1

firsts_county_tex <- firsts %>%filter(state == 'Texas') %>% group_by(county)%>% mutate((countrow = n()))
class(firsts_county_tex)
view(firsts_county_tex)

#data.frame(rbind(firsts_county_tex)) 
#class(firsts_county_tex)

plot3 <- firsts_county_tex %>%
  ggplot(aes(y= county)) + geom_histogram(stat="count")
#view(firsts_county_tex)
plot3
#viz 1

firsts %>% ggplot(aes(year)) + geom_histogram()

#viz 2

firsts %>% ggplot(aes(year)) + 
  geom_histogram()+
  facet_wrap(~ category)

#viz 3

g2 <-firsts%>% count(category, sort= TRUE) %>%
  mutate(category= fct_reorder(category,n))%>%
  ggplot(aes(n, category))+ 
  geom_col()
ggplotly(g2)



#view(firsts_joined_sf)
```

```{r}
g <- firsts%>%
  filter( state == "Texas") %>%
  ggplot(aes(year,category, 
             color = category,
             text = paste(accomplishment,":", name))) +
  geom_point() +
  #facet_wrap(~gender)+
  #scale_y_continuous()+
  labs(y= "")+
  theme( #axis.text.y = element_blank(), 
         panel.grid.major.y = element_blank(),
         axis.ticks.y = element_blank(),
         panel.grid.minor.y = element_blank(),
         legend.position = "none")+
  labs(caption =    "the firsts begin a 4 years after the 1917 roit of houston,
       Racial desegregation of the Houston Independent School District (HISD), resulting from the Civil Rights Movement in the 1950s and 1960s,[70] occurred in the 1970s")
   #g       

  ggplotly(g)
  
  describe(firsts$lat)
  

```

```{r}
plot_2 <-  firsts%>% filter(gender == 'female') %>%
  ggplot(aes(year,category, 
             color = category,
             text = paste(accomplishment,":", name))) +
  geom_point() +
  #facet_wrap(~gender)+
  #scale_y_continuous()+
  labs(y= "")+
  theme( #axis.text.y = element_blank(), 
         panel.grid.major.y = element_blank(),
         axis.ticks.y = element_blank(),
         panel.grid.minor.y = element_blank(),
         legend.position = "none")

ggplotly(plot_2)
 
 plot_7 <-  firsts%>% 
   #filter(gender == 'female') %>%
   ggplot(aes(x=year)) + geom_density()+ facet_wrap(~gender)
 plot_7

```
```{r}
West = c("California", "Oregon","Washington", "Nevada", "Idaho", "Utah", "Arizona", "Montana", "Alaska", "Hawaii")

library(dplyr)


plot_17 <-firsts%>% filter(state %in% West) %>%
  ggplot(aes(year,category, 
             color = category))+
             #text = paste(accomplishment,":", name))) +
  geom_point() +
  #facet_wrap(~gender)+
  #scale_y_continuous()+
  labs(y= "", title = " west coast")+
   xlim(1800, 2025)+
  theme( #axis.text.y = element_blank(), 
         panel.grid.major.y = element_blank(),
         axis.ticks.y = element_blank(),
         panel.grid.minor.y = element_blank(),
         legend.position = "none")

ggplotly(plot_17)



Central= c("Wyoming", "Colorado", "New Mexico", "North Dakota", "South Dakota", "Nebraska", "Kansas", "Oklahoma", "Texas", "Minnesota", "Iowa", "Missouri", "Arkansas", "Louisiana", "Wisconsin", "Illinois", "Mississippi")

plot_18 <-firsts%>% filter(state %in% Central) %>%
  ggplot(aes(year,category, 
             color = category))+
             #text = paste(accomplishment,":", name))) +
  geom_point() +
  #facet_wrap(~gender)+
  #scale_y_continuous()+
  labs(y= "", title = " Central")+
  xlim(1800, 2025)+
  theme( #axis.text.y = element_blank(), 
         panel.grid.major.y = element_blank(),
         axis.ticks.y = element_blank(),
         panel.grid.minor.y = element_blank(),
         legend.position = "none")

ggplotly(plot_18)

East= c("Michigan", "Indiana", "Kentucky", "Tennessee", "Alabama", "Ohio", "Georgia", "Florida", "South Carolina", "North Carolina", "Virginia", "West Virginia", "Delaware", "Maryland", "New Jersey", "Pennsylvania", "New York", "Connecticut", "Rhode Island", "Massachusetts", "Vermont", "New Hampshire", "Maine")

plot_19 <-firsts%>% filter(state %in% East) %>%
  ggplot(aes(year,category, 
             color = category))+
             #text = paste(accomplishment,":", name))) +
  geom_point() +
  xlim(1800, 2025)+
  #facet_wrap(~gender)+
  #scale_y_continuous()+
  labs(y= "", title = " East")+
  theme( #axis.text.y = element_blank(), 
         panel.grid.major.y = element_blank(),
         axis.ticks.y = element_blank(),
         panel.grid.minor.y = element_blank(),
         legend.position = "none")

ggplotly(plot_19)

grid.arrange(plot_17,plot_18, plot_19, ncol=1)

```




label2 <- "New York State outnumbers \n the other states"
my_colours <- c("grey80", "dodgerblue1")

states <- firsts %>% 
  filter(!is.na(state)) %>% 
  count(state) %>% 
  mutate(tot  = sum(n), pct = n/sum(n)*100,
         NY = ifelse(`state` == "New York", TRUE, FALSE)) %>%
  filter(pct > 2) %>% 
  arrange(desc(n))

states$pct <- paste(round(states$pct), "%", sep="")
  

states <- states[-c(28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43),]

ggplot(states, aes(x = n, y = reorder(state, n))) +
  geom_col() +
  theme_classic()

ggplot(states, aes(x =  n, 
                     y = reorder(state, n), fill = NY)) + 
  geom_col() + 
  scale_fill_manual(values = my_colours) + 
  # geom_text(
  #   aes(label = n, x = n - .25),
  #   colour = "red4",
  #   size = 5,
  #   nudge_x = 400,
  #   nudge_y = 0,
  #   hjust = 0.45,
  #   family="Lato") +
  geom_text(
    aes(label = pct),
    colour = "green4",
    size = 6,
    nudge_x = 1,
    nudge_y = 0,
    hjust = 0.5,
    family="Lato")+
  geom_text(
    data = data.frame(x = 25, y = 9.5, label = label2),
    aes(x = x, y = y, label = label2),
    colour="dodgerblue1",
    family="Lato",
    size = 6,
    hjust = 0.5,
    lineheight = .8,
    inherit.aes = FALSE)+
  # geom_text(
  #   data = data.frame(x = 2500, y = 0.75, label = label2),
  #   aes(x = x, y = y, label = label2),
  #   colour="green4",
  #   family="Lato",
  #   size = 4.5,
  #   hjust = 0.5,
  #   lineheight = .8,
  #   inherit.aes = FALSE)+
  geom_curve(
    data = data.frame(x = 25, y = 10, xend = 30, yend = 19.2),
    mapping = aes(x = x, y = y, xend = xend, yend = yend),
    colour = "grey15",
    size = 0.5,
    curvature = 0.25,
    arrow = arrow(length = unit(2, "mm"), type = "closed"),
    inherit.aes = FALSE) +
  labs(colour = "stoppage",
       title = "The most liberal state takes the highest number of firsts",
       subtitle = "Over time, East Coast states have gathered more firsts",
       caption = "",
       x = "Number of African American Firsts",
       y = "") +
  theme_classic()+
  theme(legend.position = "none",
        plot.title = element_text(size = 23, face = "bold", colour = "dodgerblue1"),
        plot.subtitle = element_text(size = 15),
        axis.title = element_text(size = 14),
        axis.text.x = element_text(angle = 0, hjust = 1, size = 14),
        axis.text.y = element_text(angle = 0, hjust = 1, size = 14))
        
#Due to the growth in population in the East Coast of the America being much earlier than parts fo central and western America, it can be seen from the bar chart above that the number of firsts have been significantly higher. This could be due to the migration of communities over time.
